\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\usepackage{amsmath, amssymb, amsthm}

\newtheorem{theorem}{Theorem}  
%\usepackage{unicode-math}
\usepackage{algorithm, algorithmicx}  % Para pseudocódigo
\usepackage{tikz}  % Para gráficos
\usetikzlibrary{automata, positioning}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{fullpage}

\usepackage{tcolorbox}
\tcbuselibrary{skins}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%			COMMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{todonotes}
\newcommand{\santiago}[1]{\todo[inline, color=green!20]{{\bf S:} #1}}
\newcommand{\cristian}[1]{\todo[inline, color=red!20]{{\bf C:} #1}}

\newcommand{\ox}{x\!\vdash}
\newcommand{\cx}{\dashv\!x}
\newcommand{\Vars}{\operatorname{Vars}}
\newcommand{\doc}{\operatorname{doc}}
\newcommand{\tup}{\operatorname{tup}}


\newtcolorbox{notabox}{
  colback=blue!5!white,
  colframe=blue!75!black,
  fonttitle=\bfseries,
  title=Nota,
  boxrule=1pt,
  arc=4pt,
  boxsep=5pt,
  left=6pt,
  right=6pt,
  top=6pt,
  bottom=6pt
}


\title{Equivalence between models for regex-based information extraction with errors}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We formalize and prove the equivalence between two models for regex-based information extraction with errors: (1) a semantics operating on the original document with error annotations, and (2) a semantics working on a modified document with edit operations. We demonstrate a bijective mapping between matches in both models, preserving spans and error operations under consistent index transformations. This results enable interoperability between systems using either approach.
\end{abstract}

\section{Preliminaries}

\subsection{Basic definitions}

\paragraph{Document and spans.} A document $d$ is a string over a finit alphabet $\Sigma$, we write $d= a_0a_2...a_{n-1}$ to denote a document of legth $|d| = n$. A span over a document $d$ is a pair $s = [i, j\rangle$ of natural numbers $i$ and $j$ with $0 \leq i \leq j \leq |d|$. $s$ is associated with a substring of $d$ from position $i$ to position $j-1$. We denote this substring by $d(s)$. Given two spans $s_1 = [i_1, j_1\rangle$ and $s_2 = [i_2, j_2\rangle$, the concatenation of $s_1$ and $s_2$ is defined as $s_1 \cdot s_2 = [i_1, j_2\rangle$, whenever $j_1 = i_2$.

\paragraph{Mappings.} Let $V$ be a fixed set of variables, disjoint from $\Sigma$. A \textit{mapping} is a function $\mu$ from a finite set of variables $dom(\mu) \subseteq V$ to spans. The empty mapping, denoted by $\varnothing$, is the only mapping such that $dom(\varnothing) = \varnothing$. Similarly, $[x \xrightarrow{} s]$ denotes the mapping whose domain only cointains the variable $x$, which it assign to span $s$.

\paragraph{Document spanners.} Let V be a fixed set of variables and $d \in \Sigma^*$ a document, a $\operatorname{(V,d)-tuple}$ is a mapping from V to $\operatorname{span(d)}$. A document spanner is a function that map every document $d$ to a set of $\operatorname{(V,d)-tuple}$, such that the range of each $\mu \in M$ are spans of d modeling the process of extracting information from d. For this goal, we will work with 3 models for defining spanners: RGX, ref-words and variable-set automata.



\paragraph{Regex formulas.} A \textit{regex-formula} (\textbf{RGX} for short) is a regular expression extended with variables in $V$ (called capture variables). Formally, we define the syntax with the recursive rule:

\[e \ := \ \varnothing \mid \epsilon \mid e \mid e \lor e \mid e \cdot e \mid e^* \mid x\{ \; e \; \}\]
 
\paragraph{Spanner defined by a RGX.}  
Let $R$ be a RGX and $d \in \Sigma^*$ a document.  
The semantics of $R$ is defined as a function 
$\llbracket R \rrbracket$ that assigns 
to $d$ the set of all mappings $\mu$ produced by matching $R$ against $d$:

\[
\llbracket R \rrbracket(d) := 
\{ \mu \mid \mu \text{ is obtained by a successful match of $R$ on $d$} \}.
\]

The semantics of a RGX $R$ is given recursively as follows:

\begin{itemize}

\item $\llbracket \varnothing \rrbracket(d) := \varnothing.$

\item $\llbracket \epsilon \rrbracket(d) := \{ \varnothing \}.$

\item $\llbracket R_1 \cdot R_2 \rrbracket(d) := \{ \mu_1 \cup \mu_2 \mid d = d_1 d_2,\ \mu_1 \in \llbracket R_1 \rrbracket(d_1),\ \mu_2 \in \llbracket R_2 \rrbracket(d_2),\ \operatorname{dom}(\mu_1) \cap \operatorname{dom}(\mu_2) = \varnothing \}.$

\item $\llbracket R_1 \lor R_2 \rrbracket(d) := \llbracket R_1 \rrbracket(d) \cup \llbracket R_2 \rrbracket(d).$

\item $\llbracket R^* \rrbracket(d) := \{ \mu_1 \cup \cdots \cup \mu_n \mid d = d_1 \cdots d_n,\ \mu_i \in \llbracket R \rrbracket(d_i)\ \text{para todo } i,\ \operatorname{dom}(\mu_i) \cap \operatorname{dom}(\mu_j) = \varnothing\ \text{si } i \neq j \}.$

\item $\llbracket x\{R'\} \rrbracket(d) := \{ \mu' \cup [x \mapsto s] \mid s \in \operatorname{span}(d),\ \mu' \in \llbracket R' \rrbracket(d[s]) \}.$

\end{itemize}



Thus, $\llbracket R \rrbracket(d)$ denotes the set of mappings resulting from all successful matches of $R$ on $d$.

\paragraph{Variable-set automata.} A \textit{variable-set automaton}(VA) can be understood as an $\epsilon$-NFA that is extended with capture variables in a way analogous to RGX; that is, it behaves a usual finite state automaton with $\epsilon-$transitions, except it can also open and close variables. Formally, a VA $\mathcal{A}$ is tuple $(Q, \Sigma, V, q_0, F, \delta )$, where $\Sigma$ is a finite set of alphabet symbols, V is a finite set of variables, Q is a finite set of states, $q_0 \in Q$ is an initial state, $F \subseteq Q$ is a set of final states and $\delta$ is a transition relation consisting of letter transitions of the form $(q, a, q')$ and variable transition of the form $(q, \ox, q')$ or $(q, \cx, q')$, where $q, q' \in Q$, $a \in Q$ and $x \in V$. We define the set $\Vars(\mathcal{A})$ as the set of all variables that occur in $\mathcal{A}$.

A configuration of a VA $\mathcal{A}$ over a document $d = a_0 a_1 ... a_{n-1}$ is a tuple $(q, i)$, where $q \in Q$ is the current state and $i \leq n$ is the current position in the document. A run $\rho$ of $\mathcal{A}$ is sequence of the form: \[\rho := (q_0, i_o) \xrightarrow[]{a_0}(q_1, i_1) \xrightarrow[]{a_1} \cdots \xrightarrow[]{a_{n-1}} (q_{n-1}, i_{n-1})\]where $a_i \in \Sigma \cup \{\ox, \cx\; | \; x \in V\}$ and $(q_i, a_{i+1}, q_{i+1})$. Moreover, $i_0...i_{n-1}$ is a non-decreasing sequence such that $i_0 = 1$ and $i_{n-1} = |d|$, and $i_{j+1} = i_j + 1$ if $a_{j+1} \in \Sigma$ and $i_{n-1} = i_{j}$ otherwise. 

Let $\mathcal{A} = (Q, \Sigma, V, q_0, F, \delta)$ be a VA and $d = a_0 a_1 \dots a_{n-1} \in \Sigma^*$ a document.
\paragraph{}The semantics of $\mathcal{A}$ is a spanner 
$\llbracket \mathcal{A} \rrbracket$ 
that maps $d$ to the set of mappings obtained from all accepting runs of  $d$ on $\mathcal{A}$. Each run $\rho$ induces a mapping $\mu_\rho$ defined by:
\[
\mu_\rho(x) := [i_{\ox}, i_{\cx})
\]
where $i_{\ox}$ is the position when the transition 
$\ox$ is taken, and $i_{\cx}$ is the position when 
$\cx$ is taken (assuming each $x \in V$ is opened and closed exactly once).

The semantics of $\mathcal{A}$ on $d$ is then given by:
\[
\llbracket \mathcal{A} \rrbracket(d) := 
\{ \mu_\rho \mid \rho \text{ is an accepting run of } \mathcal{A} \text{ on } d \}.
\]



\paragraph{Ref-words.} For a finite set $V$ of variables, ref-words are defined over the extended alphabet $\Sigma \cup \Gamma_V$, where $\Gamma_v := \{\ox, \cx \in V\}$, we assume that $\Gamma_V$ is disjoint with $\Sigma$. Ref-words extend strings over $\Sigma$ by encoding opening $(\ox)$ and closing $(\cx)$ of variables.
A ref-words $r \in (\Sigma \cup \Gamma_V)^*$ is valid if every ocurring variable is opened and closed exactly once. For every valid ref-word r over $(\Sigma \cup \Gamma_V)$ we define $\Vars(r)$ as the set of variables $x \in V$ which ocurr in the ref-word. More formally:
\[
\Vars(r) := \{ x \in V \; | \; \exists r^{prev}_x, r_x, r_x^{post} \in (\Sigma \cup \Gamma_V)^*.\; r = r_x^{prev} \cdot x\vdash \cdot r_x \cdot \dashv x \cdot r_x^{post}\}
\]

Intuitively, each valid ref-word encondes a mapping for some document d, the variable markers encode where the spans begin and end. Formally, we define functions $\doc$ and $\tup$ that, given a valid ref-word, output the corresponding document and mapping. The morphism $\doc: (\Sigma \cup \Gamma_V)^* \xrightarrow[]{} \Sigma^*$ is defined as:
\[
\doc(a) :=
\begin{cases}
a & \text{if } a \in \Sigma \\[6pt]
\varepsilon & \text{if } a \in \Gamma_V
\end{cases}
\]
Now, we have that every valid ref-word has a unique factorization
\[
r = r_x^{prev} \cdot x\vdash \cdot r_x \cdot \dashv x \cdot r_x^{post}
\] 
for each $x \in \Vars(r)$. We can define de function $\tup$ as 
\[
\tup(r) := \{x \xrightarrow[]{} [i_x, j_x\rangle \; | \; x \in \Vars(r), i_x = |\doc(r^{prev}_x)|, j_x = i_x + |\doc(r_x)|\}
\]
The usage of the doc morphism ensures that indices $i_x$ and $j_x$  refer to positions in the document and do not consider other variable operations. 

We also define the \emph{trace} morphism $\gamma:(\Sigma \cup \Gamma_V)^* \to \Gamma_V^*$ :

\[
\gamma(a) :=
\begin{cases}
\varepsilon & \text{if } a \in \Sigma \\[6pt]
a & \text{if } a \in \Gamma_V
\end{cases}
\]

\subsection{Edit distance}


%
%\subsection{REmatch Semantics}
%\label{sec:rematch-semantics}
%
%REmatch's formal semantics are defined through inductive rules that specify how patterns match documents and produce variable bindings. The core semantic definition appears in Table~\ref{tab:rematch-semantics}, which we explain below.
%
%\subsubsection{Semantic Foundations}
%The semantics operates on two levels:
%\begin{itemize}
%    \item $\llbracket e \rrbracket_d$: The set of all $(span, mapping)$ pairs where $e$ matches document substring $d[span]$
%    \item  $\llbracket e \rrbracket_d$: The final variable mappings produced by matching $e$ against entire document $d$
%\end{itemize}
%
%\begin{table}[htbp]
%\centering
%\caption{Inductive semantics of REQL queries}
%\label{tab:rematch-semantics}
%\begin{tabular}{ll}
%\toprule
%\textbf{Pattern} & \textbf{Semantics} \\
%\midrule
%$a$ & $\llbracket a \rrbracket_d = \{(s, \emptyset) \mid s \in \text{span}(d) \land d(s) = a\}$ \\
%$.$ & $\llbracket . \rrbracket_d = \{([i,i+1\rangle, \emptyset) \mid 0 \leq i < |d|\}$ \\
%$[w]$ & $\llbracket [w] \rrbracket_d = \{(s, \emptyset) \mid d(s) \in \text{set}(w)\}$ \\
%$\texttt{!x}\{e\}$ & $\llbracket \texttt{!x}\{e\} \rrbracket_d = \{(s, \mu) \mid \exists (s, \mu') \in \llbracket e \rrbracket_d : \mu = \mu' \cup [x \mapsto s]\}$ \\
%$e_1e_2$ & $\llbracket e_1e_2 \rrbracket_d = \{(s_1 \cdot s_2, \mu_1 \cup \mu_2) \mid (s_i, \mu_i) \in \llbracket e_i \rrbracket_d\}$ \\
%$e_1|e_2$ & $\llbracket e_1|e_2 \rrbracket_d = \llbracket e_1 \rrbracket_d \cup \llbracket e_2 \rrbracket_d$ \\
%$e^*$ & $\llbracket e^* \rrbracket_d = \bigcup_{k\geq0} \llbracket e^k \rrbracket_d$ \\
%\bottomrule
%\end{tabular}
%\end{table}

%\subsubsection{Key Semantic Rules}
%The table shows how REmatch composes matches from subexpressions:
%
%
%This semantic model enables REmatch to:
%\begin{itemize}
%    \item Track all possible matches simultaneously
%    \item Maintain correct variable binding scopes
%    \item Compose matches algebraically via span concatenation
%\end{itemize}

\subsubsection{The Edit Distance Problem}
In document processing with regular expressions, we often encounter texts with imperfections such as:
\begin{itemize}
    \item Typos (e.g., documment instead of document)
    \item Missing characters (e.g., docment)
    \item Extra insertions (e.g., docuument)


    ``a'' $\rightarrow$ ``abc''.  ins(1, b), ins(c,2) | ins (1,c) ins(1,b)


\end{itemize}

The \emph{edit distance} between two strings $D$ and $D'$, denoted by $\Delta(D,D')$, is the minimum number of elementary edit operations required to transform $D$ into $D'$. We consider three fundamental operations:\\

\subsubsection{Operations}
Given a document $D = a_1 \cdots a_n$:
\begin{center}
\begin{itemize}
    \item \textbf{Insertion}: $\mathsf{ins}(p,c)$ - Insert character $c$ at position $p$
    \[ D' = a_1 \cdots a_p c a_{p+1} \cdots a_n \]
    \item \textbf{Deletion}: $\mathsf{del}(p, c)$ - Delete character $c$ at position $p$
    \[ D' = a_1 \cdots a_{p-1} a_{p+1} \cdots a_n \]
    \item \textbf{Substitution}: $\mathsf{sub}(p,d, c)$ - Replace character $d$ with character $c$ at position $p$
    \[ D' = a_1 \cdots a_{p-1} c a_{p+1} \cdots a_n \]
\end{itemize}
\end{center}


\section{Edit Operation Application}
\label{sec:edit-application}

\subsection{Formal Definition}

Given a document $D \in \Sigma^*$ (where $D = d_1d_2\cdots d_n$) and a sequence of edit operations $E = [op_1, \ldots, op_m]$, we define the transformed document $E(D)$ recursively as:

\[E(D) = 
\begin{cases}
D & \text{if } E = \emptyset \\
E'(\mathsf{op}_1(D)) & \text{otherwise, with } E' = [op_2, op_3, \ldots, op_m]
\end{cases}
\]

where each atomic operation $\mathsf{op}_i$ modifies the document as follows:

\begin{align*}
\mathsf{ins}(p,c)(D) &= d_1\cdots d_p c d_{p+1}\cdots d_n \quad (0 \leq p \leq n) \\
\mathsf{del}(p, c)(D) &= d_1\cdots d_{p-1} d_{p+1}\cdots d_n \quad (1 \leq p \leq n) \\
\mathsf{sub}(p,d_p, c)(D) &= d_1\cdots d_{p-1} c d_{p+1}\cdots d_n \quad (1 \leq p \leq n)
\end{align*}

\subsection{Properties}

\begin{center}
	For any document $D$ and valid operation sequence $E$:
	\begin{enumerate}
		\item \textbf{Termination}: $E(D)$ always terminates in $O(|E|)$ steps
		\item \textbf{Length Change}: $|E(D)| = |D| + \#\mathsf{ins} - \#\mathsf{del}$
		\item \textbf{Order Sensitivity}: $[op_1, op_2](D) \neq [op_2, op_1](D)$ in general
	\end{enumerate}
\end{center}

\subsubsection{Example}

\begin{center}
	
	Let $D = \texttt{algorithm}$ and $E = [\mathsf{ins}(3,\texttt{h}), \mathsf{del}(6), \mathsf{sub}(2,\texttt{L})]$. Then:
	
	\begin{align*}
		E(D) &= [\mathsf{del}(6), \mathsf{sub}(2,\texttt{L})](\mathsf{ins}(3,\texttt{h})(\texttt{algorithm})) \\
		&= [\mathsf{del}(6), \mathsf{sub}(2,\texttt{L})](\texttt{alghorithm}) \\
		&= [\mathsf{sub}(2,\texttt{L})](\mathsf{del(6)}(\texttt{alghorthm})) \\
		&= [\mathsf{sub}(2,\texttt{L})](\texttt{alghothm})\\
		&= \texttt{aLgrithm} 
	\end{align*}
\end{center}


\subsection{Application of Edit Sequences on Mappings}

Let $\mu : x \to [i, j \rangle$ be a mapping and a edit operations $E = [op_1, \ldots, op_m]$, we define 
its corresponding transformation on $\mu$, denoted $E(\mu)$, as the mapping obtained by 
adjusting all affected intervals to maintain positional consistency with the edited document.

Formally, for each operation $\mathsf{op} \in E$:

The extended mapping transformation is defined recursively:
\[
E(\mu) =
\begin{cases}
	\mu & \text{if } E = \emptyset, \\[4pt]
	E'(\mathsf{op}_1(\mu)) & \text{otherwise, where } E' = [op_2, \ldots, op_m].
\end{cases}
\]

where each atomic operation $\mathsf{op}_i$ modifies the mapping as follows:

\begin{align*}
	\mathsf{ins}(p,c)(\mu) &= \begin{cases}
		[i + 1 , j+1 \rangle & p \leq i\\
		[i, j +1\rangle  & i < p \leq j\\
		[i, j \rangle   & j < p
	\end{cases}\\
	\mathsf{del}(p, c)(\mu) &= \begin{cases}
		[i - 1 , j-1 \rangle & p \leq i\\
		[i, j - 1\rangle  & i < p \leq j\\
		[i, j \rangle   & j < p
	\end{cases} \\
	\mathsf{sub}(p,d, c)(\mu) &= \mu
\end{align*}



In other words, each insertion or deletion shifts the intervals of $\mu$ that occur 
after the edit position, ensuring that $\mu$ remains consistent with the structure of 
the transformed document $E(D)$.

This definition guarantees that for any $x \in \operatorname{dom}(\mu)$,
the relative alignment between $\mu(x)$ and its corresponding content in $E(D)$ 
is preserved across arbitrary edit sequences.

\subsection{Undo edits}

\paragraph{title}Note that, given a edit sequence $E$, a document $d$ and the document resulting from applying the edit sequence over $d$, $E(d) = d'$, we can easily undo de changes taking the inverse of each operation. We will write  $E^{-1}$ the inverse sequence of edits and it is defined as follow.

For each atomic operation $op_i \in E$: 

\begin{align*}
	\mathsf{ins}(p,c)(D)^{-1} &= \mathsf{del}(p, c)(D) \\
	\mathsf{del}(p,c)(D)^{-1} &= \mathsf{ins}(p, c)(D) \\
	\mathsf{sub}(p, d, c)(D)^{-1} &= \mathsf{sub}(p, c, d)(D)
\end{align*}

And finally  $E^{-1} = [op_1^{-1}, \ldots, op_n^{-1}]$. Note that this definition works good when we want to undo the edits maded over a mapping $\mu$.
 
 
\subsection{Normalized edit sequence}

Two edit sequences $E$ and  $E'$ will we said $\textit{equivalent}$, $E \equiv E'$ if they generate identical output  $E(d) = E'(d)$ when they operate on any document $d$. Given two documents $d$ and $d'$, there exist infinte ways to transform $d$ in $d'$. In order to compare edit sequences, it is convenient to define a normalized form. For this purpose, supose that in a edit sequence $E = [op_1, \ldots, op_n]$, the atomic operation $op_i$ work in the position $d(i)$, we will write $op_i \prec op_j$ if either $i < j$

\section{Models for information extraction with errors}

Now, we will consider two diferent approaches to aboard the edit distance problem in Rematch engine. 
\subsection{Error over a modified document}

\paragraph{}The first option is the straightforward version. Take a document d, compute all the possibles documents d' that the edit distance between d and d' is less than k and prove if d' matches de REQL expresion.

This is the semantics for this approach: \\

Given a document $d \in \Sigma^*$ (where $d = d_1d_2\cdots d_n$) and a REQL expression $R$, we define the semantics of R with k error as it follows:


\[ \llbracket R^{\leq k} \rrbracket_d = \{(\mu, E) \; | \; \exists d'. E(d) = d' \land \mu \in \llbracket R \rrbracket_{d'} \land | E | \leq k  \}\]


We make a match in a new document d' that is the result of apply the edits in E over d. The span and the mapping is over d' because we only compute d' and treat it like the original document.

\subsubsection{Example}

Using the same example as before, with $d = a$, $e= a!x\{b\}c$ and $k = 2$ the following is a valid match in the new semantics:

\[([0, 3\rangle, \mu := [x \mapsto [1, 2\rangle], E = (ins(1, b), ins(2,c))\]

The document d' that make the  match is $d' = abc$, the span and mapping is over this document.

Now we give the user the new document and the list of edits that we make over the original to get the new one.

% \subsubsection{Idea para computar}

% Se puede seguir una idea similar a la que usa en el modulo de filtering, aplicar un algoritmo de light search sobre el documento original que, si es que no se llega a encontrar ningun match, pueda agregar al documento los pasos que faltaron para obtener al menos un match 


\subsection{Error over the original document}


Given a document $d \in \Sigma^*$ (where $d = d_1d_2\cdots d_n$), a REQL expression $R$, and an edit sequence $E = [op_1, \ldots, op_m]$, we define the semantics of $R$ with k errors over the \emph{original} document as:

\[
\llbracket R + k \rrbracket_d = 
\{\, (\mu, E) \;|\;
E(\mu) \in \llbracket R \rrbracket_{E(d)} \land |E| \leq k
\,\}
\]

where $\mu$ is a mapping over the original document $d$, and $E(\mu)$ denotes the mapping obtained by applying the same edit sequence $E$ to the spans of $\mu$, ensuring positional consistency with the transformed document $E(d)$.

In this formulation, the match is reported over the original document $d$, while $E$ represents the minimal sequence of edits required to make $d$ satisfy the REQL expression $R$.

%\paragraph{}Let $V$ be a finite set of variables and $\Gamma_V := \{\,\ox,\cx\mid x\in V\,\}$,
%with $\Gamma_V \cap \Sigma = \varnothing$. Given a REQL $R$, let $\mathcal{L}_{\mathrm{ref}}(R)\subseteq(\Sigma\cup\Gamma_V)^*$ be the standard ref-language of $R$,
%consisting of valid ref-words generated by $R$, the RGX $!x\{R'\}$ is equivalent to $\ox \cdot R' \cdot \cx$ wich capture a ref-word.
%We lift the $k$-error notion to ref-words as follow:

%\[
%\mathcal{L}_{\mathrm{ref}}^{\le k}(R)
%\;:=\;
%\bigl\{\, r \ \text{valid} \ \big|\ 
%\exists r' \in \mathcal{L}_{\mathrm{ref}}(R):
%\ \gamma(r)=\gamma(r') \ \land\ \Delta\bigl(\doc(r),\doc(r')\bigr)\le k \bigr\}.
%\]

%Intuitively, $r$ keeps the same sequence of variable openings/closings as some $r'$ accepted by $R$,
%while allowing up to $k$ edits on the document.

%The spanner defined by this ref-language is:

%\[\llbracket\mathcal{L}_{\mathrm{ref}}^{\le k}(R) \rrbracket_r := \{\tup(r) \mid r \in \mathcal{L}_{\mathrm{ref}}^{\le k}(R) \} \]

%\paragraph{Spanner semantics with $k$ errors.}
%For a document $d\in\Sigma^*$ and a mapping $\mu$ we can define an unique ref-word $r$ such that $\doc(r) = d$ and $\tup(r) = \mu$. The semantic of a RGX R with k errors is:
%\[
%\llbracket R \rrbracket^{\le k}(d)
%\;:=\;
 %\{(\mu, E) \mid \exists r. \doc(r) = d \land  \tup(r) = \mu \land \mu \in \llbracket\mathcal{L}_{\mathrm{ref}}^{\le k}(R) \rrbracket_r\}\]



%In both definitions, spans and mappings are taken \emph{over the original document $d$};
%edits are accounted for only in the $\Sigma$-projection used to witness compatibility with $R$.



\subsubsection{Example}


Given the document $d = a$ and the regex $e = a!x\{b\}c$, with $k = 2$ two errors available, the following is a match in this sematics.

\[
\Big([0,1\rangle,\ \mu := \{x \mapsto [1,1\rangle\},\ E = [\mathsf{ins}(1,\texttt{b}), \mathsf{ins}(1,\texttt{c})]\Big)
\]

\subsubsection{Idea para computar}

Usar un automata de levenstein que lleve registro de las posiciones del documento donde se hicieron los errores.


\section{Equivalence Proof}

\begin{theorem}
	For every document $d\in\Sigma^*$, REQL $R$, and budget $k\in\mathbb{N}$, there is a bijection between
	\[
	\llbracket R^{\le k}\rrbracket_d
	\quad\text{and}\quad
	\llbracket R + k\rrbracket_d,
	\]
	given by translating mappings via the action of the edit sequence $E$ on spans.
\end{theorem}

\begin{proof}
	We define two inverse functions between the sets and prove they are well-defined and mutually inverse.
	
\subsubsection{$\llbracket R^{\leq k} \rrbracket \implies \llbracket R + k \rrbracket_d$}
	Given $(\mu',E)\in \llbracket R^{\le k}\rrbracket_d$, then there exists $d'$ with $E(d)=d'$ and $\mu'\in\llbracket R\rrbracket_{d'}$ and $|E|\le k$.
	Let $\mu := E^{-1}(\mu')$. Then $E(\mu)=E(E^{-1}(\mu'))=\mu'$, hence
	\[
	E(\mu)\in \llbracket R\rrbracket_{E(d)}.
	\]
	By definition of the original-document semantics, $(\mu,E)\in \llbracket R + k\rrbracket_d$ with $|E|\le k$.
	

\subsubsection{$\llbracket R + k \rrbracket_d \implies \llbracket R ^{\leq k} \rrbracket_d$}	
	Given $(\mu,E)\in \llbracket R + k\rrbracket_d$, then $E(\mu)\in\llbracket R\rrbracket_{E(d)}$ and $|E|\le k$.
	Let $d' = E(d)$. Then $(E(\mu),E)\in \llbracket R^{\le k}\rrbracket_d$ by the modified-document semantics.
	
	
\end{proof}




\end{document}