%!TEX root = ../main/main.tex

We now introduce the semantics for information extraction with errors. 
Let $S$ be a document spanner. 
Given a document $d$ and an error budget $k$, we consider edit scripts $E$ with $|E|\leq k$ and evaluate $S$ over the edited document $E(d)$. 

\subsection{Approximate extraction semantics}

The semantics evaluates the spanner over an edited version of the input document, but the resulting mappings are transported backwards through the edit script to refer to valid spans within the original input document.

\begin{definition}
	Let $S$ be a document spanner, let $d \in \Sigma^*$, and let $k \in \mathbb{N}$. 
	The \emph{approximate semantics} of $S$ over $d$ with error budget $k$ is defined as
	\[
	\editsem{S}{k}{d}
	=
	\left\{
	(E^{-1}(\mu),E)
	\;\middle|\;
	E \in \operatorname{Edits}^*,\
	|E| \leq k,
	\text{ and }
	\mu \in \llbracket S \rrbracket (E(d))
	\right\}.
	\]
\end{definition}

In this semantics, the script $E$ witnesses how the input document is modified to satisfy the spanner. While the mapping $\mu$ is computed directly on the resulting edited document $E(d)$, the returned mapping $E^{-1}(\mu)$ correctly refers to the coordinates of the original document $d$. 


\subsubsection{Examples}

To illustrate how the inverse transport adjusts the spans back to the original document, we present two scenarios: an insertion and a deletion occurring inside an extracted span.

\noindent
\textbf{Insertion inside an extracted span.}

Consider the original document $d=\mathtt{ac}$ and the edit script $E=\ins_{2,\mathtt{b}}$.
Applying the edit yields the document $E(d)=\mathtt{abc}$.
Suppose the spanner extracts the following mapping over the edited document:
\[
\mu =
[
x \mapsto \Span{1,4},
y \mapsto \Span{2,3}
].
\]
This corresponds to the extraction $x\{\mathtt{a}\,y\{\mathtt{b}\}\,\mathtt{c}\}$ over $\mathtt{abc}$.

Since the edit inserted $\mathtt{b}$ at position $2$, the inverse transport $E^{-1}$ acts as a deletion at position $2$. This reduces by one the upper limit of the spans that cross this position. 
Thus, transporting the mapping back to the original document yields:
\[
E^{-1}(\mu)
=
[
x \mapsto \Span{1,3},
y \mapsto \Span{2,2}
].
\]
In the coordinates of the original document $d=\mathtt{ac}$, the variable $x$ captures the entire document, while $y$ becomes an empty span between $\mathtt{a}$ and $\mathtt{c}$. We can visualize this effect as:
\[
\text{edited: } x\{\mathtt{a}\,y\{\mathtt{b}\}\,\mathtt{c}\}
\qquad
\leadsto
\qquad
\text{original: } x\{\mathtt{a}\,y\{\epsilon\}\,\mathtt{c}\}.
\]


\noindent
\textbf{Deletion inside an extracted span.}

Consider the original document $d=\mathtt{abbc}$ and the edit script $E=\del_2$.
Applying this deletion results in the edited document $E(d)=\mathtt{abc}$.
Assume the spanner extracts the mapping $\mu$ over $E(d)$ given by:
\[
\mu =
[
x \mapsto \Span{1,4},
y \mapsto \Span{3,4}
].
\]
This corresponds to the extraction $x\{\mathtt{a}\mathtt{b}\,y\{\mathtt{c}\}\}$ over $\mathtt{abc}$.

Because the original edit script deleted the symbol at position $2$ of $d$, the inverse transport $E^{-1}$ behaves as an insertion at position $2$. This effectively shifts the coordinates of the spans that occur after the deletion one position to the right, aligning them with the original text. 
Therefore, the inverse mapping is:
\[
E^{-1}(\mu)
=
[
x \mapsto \Span{1,5},
y \mapsto \Span{4,5}
].
\]
Over the original document $d=\mathtt{abbc}$, the variable $x$ correctly captures the entire string $\Span{1,5}$, and $y$ captures the final $\mathtt{c}$ at $\Span{4,5}$.

%!TEX root = ../main/main.tex

\subsection{Validity of the Approximate Semantics}

To ensure that $\llbracket S_{\leq k} \rrbracket$ acts as a valid extraction framework over the original document, we must prove that any mapping transported back through $E^{-1}$ only contains valid spans over $d$. We prove this by induction on the sequence of edit operations.

First, we show that a single inverse edit operation preserves span validity with respect to the pre-edited document.

\begin{lemma}\label{lemma:single_edit_validity}
	Let $d \in \Sigma^*$ be a document and $e \in \operatorname{Edits}$ an atomic edit operation. 
	If $s = \Span{i,j} \in \operatorname{span}(e(d))$, then $e^{-1}(s) \in \operatorname{span}(d)$.
\end{lemma}

\begin{proof}
	Let $n = |d|$. We analyze the three possible types of atomic edits applied to $d$ to verify that $e^{-1}(s)$ is a valid span over $d$. This means its transformed coordinates must fall strictly within the range $[1, n+1]$ and preserve the order $i \leq j$.
	
	\begin{itemize}
		\item \textbf{$e = \sub_{p,c}$:} 
		The length of the document does not change, so $|e(d)| = n$. 
		Thus, $s = \Span{i,j} \in \operatorname{span}(e(d))$ implies $1 \leq i \leq j \leq n+1$. 
		Since $e^{-1} = \sub_{p,c}$, the inverse transport leaves the span completely unchanged: $e^{-1}(s) = \Span{i,j}$. 
		Both bounds trivially remain valid over $d$.
		
		\vspace{0.3cm}
		
		\item \textbf{$e = \ins_{p,c}$:} 
		The edited document grows by one character, so $|e(d)| = n+1$. 
		Thus, $s = \Span{i,j} \in \operatorname{span}(e(d))$ implies $1 \leq i \leq j \leq n+2$. 
		A valid insertion on the original document $d$ occurs at position $1 \leq p \leq n+1$. The inverse transport is $e^{-1} = \del_p$. According to its definition, we analyze three sub-cases:
		\begin{itemize}
			\item \textit{Case $p < i \leq j$:} Both coordinates are strictly greater than $p$, so both decrease by one. The new upper bound is $j-1 \leq (n+2)-1 = n+1$. The new lower bound is $i-1 \geq p \geq 1$.
			\item \textit{Case $i \leq p < j$:} Only $j$ is strictly greater than $p$, so only $j$ decreases. The new upper bound becomes $j-1 \leq n+1$. The lower bound remains $i \geq 1$.
			\item \textit{Case $i \leq j \leq p$:} Neither coordinate changes. Since $p$ is bounded by the maximum valid insertion position on $d$ ($p \leq n+1$), we inherently have $j \leq p \leq n+1$. The lower bound remains $i \geq 1$.
		\end{itemize}
		In all three sub-cases, the new coordinates are bounded within $[1, n+1]$, wich mean $e^{-1}(s) \in \operatorname{span}(d)$.
		
		\vspace{0.3cm}
		\item \textbf{$e = \del_p$:} 
		The edited document shrinks by one character, so $|e(d)| = n-1$. 
		Thus, $s = \Span{i,j} \in \operatorname{span}(e(d))$ implies $1 \leq i \leq j \leq n$. 
		A valid deletion on the original document $d$ occurs at position $1 \leq p \leq n$. The inverse transport is $e^{-1} = \ins_{p,c}$. We again analyze three sub-cases:
		\begin{itemize}
			\item \textit{Case $p \leq i \leq j$:} Both coordinates are greater than or equal to $p$, so both increase by one. The new upper bound is $j+1 \leq n+1$. The new lower bound is $i+1 \geq p+1 \geq 2 > 1$.
			\item \textit{Case $i < p \leq j$:} Only $j$ increases. The new upper bound is $j+1 \leq n+1$. The lower bound remains $i \geq 1$.
			\item \textit{Case $i \leq j < p$:} Neither coordinate changes. Since $s$ was a valid span over $e(d)$, we know $j \leq n < n+1$. The lower bound remains $i \geq 1$.
		\end{itemize}
		In all three sub-cases, the new coordinates fall within $[1, n+1]$.
	\end{itemize}
	Since the limits are respected and the monotonic nature of the inverse operations preserves $i \leq j$, we conclude that $e^{-1}(s) \in \operatorname{span}(d)$.
\end{proof}

Now we extend this property to arbitrary edit scripts, proving that the returned mappings belong to the original document.

\begin{proposition}
	Let $d \in \Sigma^*$ be a document, $E = e_1e_2\cdots e_m \in \operatorname{Edits}^*$ an edit script, and $\mu$ a mapping over $E(d)$. 
	Then $E^{-1}(\mu)$ is a valid mapping over $d$.
\end{proposition}

\begin{proof}
	We proceed by induction on the length of the edit script, $m = |E|$.
	
	\textbf{Base case ($m=0$):} 
	If $E = \epsilon$, then $E(d) = d$. 
	The inverse transport is $\epsilon^{-1}(\mu) = \mu$. 
	Since $\mu$ is a mapping over $E(d)=d$, it is trivially a valid mapping over $d$.
	
	\textbf{Inductive step:} 
	Assume the proposition holds for all edit scripts of length $m-1$.
	We can decompose $E$ as $E = E' e_m$, where $E' = e_1 \cdots e_{m-1}$. 

	The edited document is $E(d) = e_m(E'(d))$, so $\mu \in \operatorname{e_m(E'(d))}$, the Lemma~\ref{lemma:single_edit_validity} guarantees that $e_m^{-1}(\mu)$ is a valid mapping over the intermediate document $E'(d)$.
	
	Since $e_m^{-1}(\mu)$ is a valid mapping over $E'(d)$ and $|E'| = m-1$, we apply the inductive hypothesis. 
	It follows that $(E')^{-1}(e_m^{-1}(\mu))$ is a valid mapping over $d$ and we have $E^{-1}(\mu) = (E')^{-1}(e_m(\mu))$. We conclude that $E^{-1}(\mu)$ is a valid mapping over $d$.
\end{proof}

Since every tuple in our semantics applies exactly this inverse transport, the approximate semantics correctly projects its extractions as valid mappings over the original document $d$.